---
title: "Annotation Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotation Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
on_ci <- identical(Sys.getenv("NOT_CRAN"), "true")
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  eval       = on_ci,
  cache      = on_ci,
  cache.path = "cache/annotation-workflow-",
  message    = FALSE,
  warning    = FALSE,
  error      = TRUE
)
```

## Overview

Once you have extracted DOIs, ISBNs, or PMIDs from Wikipedia wikitext, the
next step is to enrich them with bibliographic metadata, citation counts, and
social-media attention scores.  **wikilite** provides four annotation pathways:

| Database | DOI | ISBN | Function |
|----------|-----|------|----------|
| EuropePMC | ✓ | — | `annotate_doi_list_europmc()` |
| CrossRef | ✓ | — | `annotate_doi_list_cross_ref()`, `annotate_doi_to_bibtex_cross_ref()` |
| Altmetric | ✓ | ✓ | `annotate_doi_list_altmetrics()`, `annotate_isbn_list_altmetrics()` |
| Google Books | — | ✓ | `annotate_isbn_google()` |
| Open Library | — | ✓ | `annotate_isbn_openlib()` |

## Setup

```{r setup-pkg}
library(wikilite)

# Fetch the most recent revision of a test article
recent <- get_article_most_recent_table("Zeitgeber")
text   <- recent$`*`
```

## Extract identifiers

```{r extract-ids}
# DOIs
doi_df   <- get_regex_citations_in_wiki_table(recent, pkg.env$doi_regexp)
doi_list <- unique(doi_df$citation_fetched)
cat("Found", length(doi_list), "unique DOIs\n")

# ISBNs
isbn_df   <- get_regex_citations_in_wiki_table(recent, pkg.env$isbn_regexp)
isbn_list <- unique(isbn_df$citation_fetched)
cat("Found", length(isbn_list), "unique ISBNs\n")
```

## Annotate DOIs with EuropePMC

EuropePMC is open-access, returns rich open-access status and citation counts,
and is the recommended first-pass annotation service.

```{r epmc}
# Annotate the first 5 DOIs
epmc_results <- annotate_doi_list_europmc(doi_list[1:5])

# Useful columns: title, authorString, journalTitle, pubYear, citedByCount
head(epmc_results[, c("doi", "title", "journalTitle", "pubYear",
                      "citedByCount", "isOpenAccess")])
```

## Annotate DOIs with CrossRef

CrossRef provides comprehensive bibliographic metadata and is the authoritative
source for DOI resolution.

```{r crossref-meta}
crossref_results <- annotate_doi_list_cross_ref(doi_list[1:5])
head(crossref_results)
```

Retrieve BibTeX entries directly:

```{r crossref-bib}
bibtex_entries <- annotate_doi_to_bibtex_cross_ref(doi_list[1:3])
cat(bibtex_entries[[1]])
```

Export to a `.bib` file:

```{r bib-export}
export_doi_to_bib(doi_list[1:10], file_name = "zeitgeber_refs.bib")
```

## Altmetric scores for DOIs

Altmetric tracks how often a paper is mentioned on social media, in news
articles, policy documents, and patents.

```{r altmetric-doi}
# Note: doi_list must be passed as a list, not a character vector
alt_results <- annotate_doi_list_altmetrics(list(doi_list[1:5]))

# Key columns
if (nrow(alt_results) > 0) {
  cols <- intersect(c("title", "doi", "score",
                      "cited_by_tweeters_count",
                      "cited_by_posts_count"), names(alt_results))
  head(alt_results[, cols])
}
```

## Identify the top-cited papers across multiple articles

`get_top_cited_wiki_papers()` combines EuropePMC annotation and CrossRef
citation counts to rank the most impactful papers cited across your article
set.

```{r top-cited}
articles  <- c("Zeitgeber", "Advanced sleep phase disorder", "Sleep deprivation")
recent_all <- get_category_articles_most_recent(articles)
doi_all    <- get_regex_citations_in_wiki_table(recent_all, pkg.env$doi_regexp)

top40 <- get_top_cited_wiki_papers(doi_all)
head(top40[, c("title", "journalTitle", "pubYear",
               "wiki_count", "count", "cited_in_wiki_art")])
```

## Annotate ISBNs

### Google Books

Google Books returns metadata for most widely published books.

```{r google-books}
# Single ISBN
book_meta <- annotate_isbn_google("9780156031356")
book_meta[, c("title", "publisher", "publishedDate", "categories")]
```

### Open Library

Open Library is an alternative book database.

```{r open-library}
book_meta_ol <- annotate_isbn_openlib("9780156031356")
str(book_meta_ol)
```

### Altmetric scores for ISBNs

```{r altmetric-isbn}
isbn_alt <- annotate_isbn_list_altmetrics(list(isbn_list[1:3]))
head(isbn_alt)
```

## Revert-trend analysis

`get_revert_counts()` queries all Wikipedia revisions in a time window and
returns articles ranked by their number of revert-tagged edits (undo /
rollback).  This is useful for identifying disputed or frequently vandalized
articles.

```{r reverts}
# One-hour window on 12 December 2018
reverts <- get_revert_counts(
  start = "20181212010000",
  end   = "20181212000000"
)

# Top 10 most reverted articles in that hour
head(reverts, 10)
```

## Putting it all together

A typical full workflow:

```{r full-workflow}
# 1. Define your article list
articles <- get_pagename_in_cat("Circadian rhythm")

# 2. Fetch the most recent revision of each article
recent <- get_category_articles_most_recent(articles)

# 3. Extract DOIs
doi_df   <- get_regex_citations_in_wiki_table(recent, pkg.env$doi_regexp)
doi_list <- unique(doi_df$citation_fetched)

# 4. Annotate via EuropePMC
annotated <- annotate_doi_list_europmc(doi_list)

# 5. Find the top-cited papers
top_papers <- get_top_cited_wiki_papers(doi_df)

# 6. Export
openxlsx::write.xlsx(annotated,   "circadian_doi_annotated.xlsx")
openxlsx::write.xlsx(top_papers,  "circadian_top_papers.xlsx")
export_doi_to_bib(annotated$doi,  "circadian_refs.bib")
```
