---
title: "Citation Analysis with wikilite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Citation Analysis with wikilite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
on_ci <- identical(Sys.getenv("NOT_CRAN"), "true")
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  eval       = on_ci,
  cache      = on_ci,
  cache.path = "cache/citation-analysis-",
  message    = FALSE,
  warning    = FALSE,
  error      = TRUE
)
```

## Introduction

This vignette demonstrates how to extract, count, parse, and visualise
citations from Wikipedia articles using **wikilite**.  We will work with a
small set of articles about sleep science to illustrate each step.

## Fetch article data

```{r fetch}
library(wikilite)

articles <- c("Zeitgeber",
              "Advanced sleep phase disorder",
              "Sleep deprivation",
              "Circadian rhythm")

# Most recent revision of each article
recent <- get_category_articles_most_recent(articles)
```

## Count citations using built-in patterns

The package stores all built-in regular expressions in the `pkg.env`
environment.  You can inspect them:

```{r regexp-list}
names(pkg.env$regexp_list)
```

Apply a single pattern:

```{r single-pattern}
doi_regexp <- pkg.env$doi_regexp
dois <- get_regex_citations_in_wiki_table(recent, doi_regexp)
head(dois)
```

Apply every pattern at once and receive a named list:

```{r all-patterns}
all_results <- extract_citations_regexp(recent)

# Each element is a data frame with art, revid, citation_fetched
sapply(all_results, nrow)
```

## Count individual elements per article

```{r per-article-counts}
results <- lapply(seq_len(nrow(recent)), function(i) {
  text <- recent$`*`[i]
  data.frame(
    art          = recent$art[i],
    n_doi        = get_doi_count(text),
    n_isbn       = get_ISBN_count(text),
    n_url        = get_urlCount(text),
    n_ref        = get_refCount(text),
    n_hyperlinks = get_hyperlinkCount(text),
    sci_score    = get_sci_score(text),
    sci_score2   = get_sci_score2(text)
  )
})

count_table <- do.call(rbind, results)
count_table
```

## Parse Citation Style 1 templates

`parse_article_ALL_citations()` splits every CS1 template into a tidy long
data frame where each row is one field:

```{r parse-one}
# For a single article
zeitgeber_recent <- recent[recent$art == "Zeitgeber", ]
parsed_one <- parse_article_ALL_citations(zeitgeber_recent$`*`)
head(parsed_one)
```

Parse all articles at once:

```{r parse-all}
parsed <- get_parsed_citations(recent)
# Columns: art, revid, type, id_cite, variable, value
head(parsed)
```

## Summarise citation types

```{r cite-type}
cite_types <- get_citation_type(recent)
head(cite_types)

# Distribution of journal / web / news / book citations
plot_distribution_source_type(cite_types)
```

## Top sources

```{r top-sources}
# Top 20 journal titles across all articles
plot_top_source(parsed, "journal")

# Top 20 publishers
plot_top_source(parsed, "publisher")

# Generate charts for multiple fields at once
get_pdfs_top20source(parsed,
  source_types_list = c("journal", "publisher", "author"))
```

## SciScore in depth

**SciScore** measures the proportion of CS1 citations that use the
`cite journal` template — a proxy for how scientifically sourced an article
is.

```{r sciscore-loop}
scores <- sapply(seq_len(nrow(recent)), function(i) {
  get_sci_score(recent$`*`[i])
})
names(scores) <- recent$art
sort(scores, decreasing = TRUE)
```

**SciScore2** divides the count of DOIs by the total number of `<ref>` tags.
A value near 1 suggests that almost every reference cites a publication with
a DOI.

```{r sciscore2-loop}
scores2 <- sapply(seq_len(nrow(recent)), function(i) {
  get_sci_score2(recent$`*`[i])
})
names(scores2) <- recent$art
scores2
```

## Identify the most-cited scientific papers

`get_top_cited_wiki_papers()` finds the 40 most frequently cited DOIs across
all supplied articles, annotates them via EuropePMC and CrossRef, and adds
per-article citation counts.

```{r top-cited}
doi_df <- get_regex_citations_in_wiki_table(recent, pkg.env$doi_regexp)
top_papers <- get_top_cited_wiki_papers(doi_df)
head(top_papers[, c("title", "journalTitle", "pubYear",
                    "wiki_count", "cited_in_wiki_art")])
```

## Explore article revision history

For longitudinal analysis, retrieve the full revision history:

```{r history}
history <- get_article_full_history_table("Zeitgeber",
                                          date_an = "2022-01-01T00:00:00Z")

# Extract DOIs from every revision and track when they first appeared
doi_history <- get_regex_citations_in_wiki_table(history, pkg.env$doi_regexp)

# How many unique DOIs appear by revision?
length(unique(doi_history$citation_fetched))
```

## Export results

```{r export}
# Save the parsed citation table to Excel
openxlsx::write.xlsx(parsed, "sleep_articles_parsed_citations.xlsx")

# Export the full doi count table
openxlsx::write.xlsx(count_table, "sleep_articles_counts.xlsx")

# Export all regex matches to separate xlsx files
export_extracted_citations_xlsx(recent, "sleep_articles")
```
