---
title: "Introduction to wikilite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to wikilite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
on_ci <- identical(Sys.getenv("NOT_CRAN"), "true")
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = on_ci,
  cache     = on_ci,
  cache.path = "cache/introduction-",
  message   = FALSE,
  warning   = FALSE,
  error     = TRUE
)
```

## Overview

**wikilite** is an R toolkit for mining Wikipedia article revision history and
citations.  It connects to the public MediaWiki API to retrieve structured data
frames that are ready for downstream analysis, annotation, and visualisation.

Key capabilities:

| Area | Key functions |
|------|---------------|
| History retrieval | `get_article_full_history_table()`, `get_article_most_recent_table()` |
| Category navigation | `get_pagename_in_cat()`, `get_subcat_with_depth()` |
| Citation counting | `get_doi_count()`, `get_refCount()`, `get_ISBN_count()` |
| Citation parsing | `parse_article_ALL_citations()`, `get_parsed_citations()` |
| Quality metrics | `get_sci_score()`, `get_sci_score2()` |
| Annotation | `annotate_doi_list_europmc()`, `annotate_doi_list_cross_ref()` |
| Revert trends | `get_revert_counts()` |
| Visualisation | `plot_article_creation_per_year()`, `page_edit_plot()` |

## Installation

```{r install, eval = FALSE}
# CRAN (once available)
install.packages("wikilite")

# Development version from GitHub
remotes::install_github("jsobel1/wikilite")
```

## Quick-start: a single article

```{r single-article}
library(wikilite)

# Most recent revision of the Zeitgeber article
recent <- get_article_most_recent_table("Zeitgeber")

# The data frame has columns: art, revid, user, timestamp, size, comment, *
# The "*" column contains the full raw wikitext.
names(recent)
```

Count citations directly from the wikitext:

```{r counts}
text <- recent$`*`

get_doi_count(text)        # number of DOIs
get_refCount(text)         # number of <ref>...</ref> blocks
get_ISBN_count(text)       # number of ISBNs
get_hyperlinkCount(text)   # number of [[...]] links
get_sci_score(text)        # fraction of citations that are journal citations
```

## Multiple articles via a category

```{r category}
# List all articles in a Wikipedia category
sleep_articles <- get_pagename_in_cat("Circadian rhythm")
head(sleep_articles)

# Fetch the most recent revision for each article
sleep_recent <- get_category_articles_most_recent(sleep_articles)
```

## Extract DOIs

```{r doi-extract}
doi_regexp <- pkg.env$doi_regexp   # "10\\.\\d{4,9}/[-._;()/:a-z0-9A-Z]+"

dois <- get_regex_citations_in_wiki_table(sleep_recent, doi_regexp)
head(dois)
```

## Visualise article creation over time

```{r viz-creation}
initial <- get_category_articles_creation(sleep_articles)

# Cumulative count (default)
plot_article_creation_per_year(initial, name_title = "Circadian rhythm articles")

# Annual counts
plot_article_creation_per_year(initial, name_title = "Annual creation",
                               Cumsum = FALSE)
```

## Static article timeline

```{r viz-timeline}
plot_static_timeline(initial)
```

## Page-view and edit-activity plots

```{r viz-activity}
# Daily page views
page_view_plot("Zeitgeber", start = "2020010100", end = "2021010100")

# Weekly edits
page_edit_plot("Zeitgeber", start = "2020010100", end = "2021010100")
```

## Export to Excel and BibTeX

```{r export}
# Revision table → Excel
write_wiki_history_to_xlsx(recent, "zeitgeber")

# All regex matches → one xlsx file per pattern
export_extracted_citations_xlsx(sleep_recent, "sleep_articles")

# DOIs → BibTeX
export_doi_to_bib(unique(dois$citation_fetched)[1:10], "references.bib")
```

## Next steps

* **Citation Analysis** vignette — deep dive into CS1 template parsing,
  SciScore, and top-cited paper identification.
* **Annotation Workflow** vignette — enrich extracted DOIs and ISBNs with
  metadata from EuropePMC, CrossRef, Altmetric, Google Books, and Open Library.
