---
title: "Multi-Article Analysis and Interactive Networks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-Article Analysis and Interactive Networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
on_ci <- identical(Sys.getenv("NOT_CRAN"), "true")
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  eval       = on_ci,
  cache      = on_ci,
  cache.path = "cache/multi-article-networks-",
  message    = FALSE,
  warning    = FALSE,
  error      = TRUE
)
```

## Overview

**wikilite** includes a suite of interactive visualisations for comparing
multiple Wikipedia articles simultaneously.  This vignette covers:

1. Building article lists from category pages
2. The interactive Gantt-style timeline
3. Three network visualisations (co-citation, publication, wikilink)
4. Citation latency analysis across articles

All network functions require **visNetwork** (`install.packages("visNetwork")`).

## 1. Build an article list

You can supply article titles directly or derive them from a Wikipedia category.

```{r article-list}
library(wikilite)

# Option A — explicit titles
articles <- c(
  "Zeitgeber",
  "Circadian rhythm",
  "Sleep deprivation",
  "Advanced sleep phase disorder",
  "Non-24-hour sleep-wake disorder"
)

# Option B — browse a category
sleep_articles <- get_pagename_in_cat("Circadian rhythm")
head(sleep_articles)

# Option C — explore subcategories first
subcats <- get_subcat_table("Chronobiology")
head(subcats)

# Then pull articles from a chosen subcategory
sub_articles <- get_pagename_in_cat("Sleep disorders")
```

## 2. Interactive Gantt timeline

`plot_interactive_timeline()` shows each article as a horizontal bar spanning
from its creation date to its last edit, coloured by a chosen metric.

```{r timeline}
# Default: colour bars by SciScore
plot_interactive_timeline(articles)

# Colour by current article size instead
plot_interactive_timeline(articles, color_by = "size")

# No colouring — uniform colour for clean presentation
plot_interactive_timeline(articles, color_by = "none")

# Restrict to a historical snapshot
plot_interactive_timeline(articles, date_an = "2021-01-01T00:00:00Z")

# French Wikipedia
plot_interactive_timeline(
  c("COVID-19", "Syndrome respiratoire aigu sévère"),
  lang = "fr"
)
```

Hovering over a bar shows the article name, creation date, last-edit date,
first editor, and SciScore.  Clicking a dot opens the article on Wikipedia.

## 3. Co-citation network

Articles sharing at least `min_shared_dois` DOIs are connected by an edge.
Node size scales with the number of shared citations; thicker edges indicate
more overlap.

```{r cocitation}
# Articles sharing ≥ 1 DOI (default)
plot_article_cocitation_network(articles)

# Require ≥ 3 shared DOIs for an edge (cleaner graph)
plot_article_cocitation_network(articles, min_shared_dois = 3)
```

The network is interactive: drag nodes, zoom, and click any node to open
the Wikipedia article in a new tab.

## 4. Publication bipartite network

A bipartite (two-mode) network where **article** nodes (blue boxes) connect
to **DOI** nodes (orange dots).  Node size indicates how many articles cite
each DOI, making highly shared papers immediately visible.

```{r pubnet}
# Show papers cited by ≥ 2 articles (default)
plot_article_publication_network(articles)

# Show up to 80 top DOIs, require at least 3 citing articles
tryCatch(
  plot_article_publication_network(
    articles,
    top_n_dois     = 80L,
    min_wiki_count = 3L
  ),
  error = function(e) message("Note: ", conditionMessage(e))
)

# Enrich paper nodes with EuropePMC metadata (title, journal, year)
plot_article_publication_network(articles, annotate = TRUE)
```

DOI nodes link directly to `doi.org` when clicked.

## 5. Wikilink network

Shows how articles in your set link **to** each other.  Directed edges point
from the linking article to the linked article.  By default, only links
between articles in your list are shown (`only_internal = TRUE`).

```{r wikilink}
# Only links between articles in the list
plot_article_wikilink_network(articles)

# Also include outgoing links to other Wikipedia articles
plot_article_wikilink_network(articles,
                               only_internal = FALSE,
                               top_n_links   = 60L)
```

## 6. Citation latency analysis

How long after a paper is published does it take to appear in Wikipedia?
Use `compute_citation_latency()` together with EuropePMC annotation.

```{r latency}
# Fetch most recent revisions
recent  <- get_category_articles_most_recent(articles)

# Extract DOIs
doi_df  <- get_regex_citations_in_wiki_table(recent, pkg.env$doi_regexp)

# Annotate with EuropePMC publication dates
epmc_df <- annotate_doi_list_europmc(unique(doi_df$citation_fetched))

# Compute latency (days from publication → Wikipedia insertion)
latency <- compute_citation_latency(doi_df, epmc_df)

# Overall latency distribution
plot_latency_distribution(latency)

# Stratified: preprints vs. journal articles (includes KS-test p-value)
plot_latency_distribution(latency, stratify_by = "is_preprint")

# Segment plot: each DOI's gap as a horizontal line
get_segment_history_doi_plot(latency, "Zeitgeber")

# Scatter: insertion date vs. latency
get_dotplot_history(latency, "Zeitgeber")
```

Key columns in the latency data frame:

| Column | Description |
|--------|-------------|
| `citation_fetched` | DOI |
| `firstPublicationDate` | EuropePMC publication date |
| `latency_days` | Days from publication to Wikipedia insertion |
| `is_preprint` | `TRUE` for bioRxiv/medRxiv DOIs (`10.1101/…`) |

## 7. Batch workflow example

The following code replicates a typical research workflow: browse a category,
download recent revisions, extract citations, annotate, and visualise.

```{r full-example}
library(wikilite)

# 1. Browse the category tree
subcats <- get_subcat_table("Chronobiology")

# 2. Choose a category and list its articles
articles <- get_pagename_in_cat("Circadian rhythm")

# 3. Visualise the timeline
plot_interactive_timeline(articles, color_by = "sciscore")

# 4. Explore citation sharing
plot_article_cocitation_network(articles, min_shared_dois = 2)
plot_article_publication_network(articles, annotate = TRUE)

# 5. Latency analysis
recent  <- get_category_articles_most_recent(articles)
doi_df  <- get_regex_citations_in_wiki_table(recent, pkg.env$doi_regexp)
epmc_df <- annotate_doi_list_europmc(unique(doi_df$citation_fetched))
latency <- compute_citation_latency(doi_df, epmc_df)
plot_latency_distribution(latency, stratify_by = "is_preprint")

# 6. Export
openxlsx::write.xlsx(latency, "circadian_latency.xlsx")
export_doi_to_bib(unique(doi_df$citation_fetched), "circadian_refs.bib")
```

## 8. Multilingual analysis

All functions accept a `lang` parameter for any Wikipedia language edition.

```{r multilingual}
# Compare sleep-science articles in French Wikipedia
fr_articles <- get_pagename_in_cat("Rythme circadien", lang = "fr")
plot_interactive_timeline(fr_articles, lang = "fr", color_by = "sciscore")

# Co-citation network in German Wikipedia
de_articles <- get_pagename_in_cat("Chronobiologie", lang = "de")
plot_article_cocitation_network(de_articles, lang = "de", min_shared_dois = 2)
```

## 9. Troubleshooting

| Problem | Solution |
|---------|----------|
| `visNetwork` not available | `install.packages("visNetwork")` |
| "No DOIs found" in network | Lower `min_wiki_count` or `min_shared_dois` |
| Network is very dense | Increase `min_shared_dois` or `top_n_dois` |
| Slow for large article sets | Use `date_an` to restrict to a historical snapshot |
| `EuropePMC annotation failed` | Check internet access; try again later |
